Evaluating computational methods for modeling off-normal operation of gas centrifuge cascades

### 2018-06-28 dis_clusion.tex
finish with gonuke comments

It is really interesting to observe that when the cascade is left completely
untouched (Model A) or when it is slightly retuned to maintain the tails to
product enrichment factor as well as the cut of each centrifuges (Model C),
chaining the cascade we can observe high increase of the enrichment at each
level.  On the contrary, when retuning the cut of each centrifuges to maintain
the ideal state of the cascades (Model B) while chaining them, the \gls{HEU}
production rate is favored over the enrichment gain.

The tails recycling allows for each model a huge gain in productivity, even for
then model B where the number of levels required to reach $90w\%$ of $^{235}$U
in the uranium does not change. Even if no cascades chaining allows to retrieve
the same production rate as a direct enrichment, the model B with tail
reprocessing reach about $80\%$ of an optimum production, which is far from
being negligible, allowing to build up a Significant Quantity of \gls{HEU} in
less than 8 months\ldots

>>>>

We can observe that when the cascade is left completely untouched (Model A) or
when it is slightly retuned to maintain the tails to product enrichment factor
as well as the cut of each centrifuges (Model C), chaining the cascade we can
observe high increase of the enrichment at each level.  On the contrary, when
retuning the cut of each centrifuges to maintain the ideal state of the cascades
(Model B) while chaining them, the \gls{HEU} production rate is favored over the
enrichment gain.

The tails recycling allows each model to achieve a large gain in productivity,
even for then model B in which the number of levels required to reach $90w\%$ of
$^{235}$U in the uranium does not change. Even if no cascade chaining options
achieves the same production rate as direct enrichment, the model B with tail
recycling reached about $80\%$ of the optimum production rate. Such production
rate would allow the build up of a Significant Quantity of \gls{HEU} in less
than 8 months\ldots

### 2018-06-28 experiement.tex
finish with gonuke comments

\subsubsection{reference}

@@

\subsubsection{default cascade}

The default cascade is the cascade design for normal civilian enrichment
operation, enriching natural uranium to about $3.5w\%$, with a tails assay close
to $0.3w\%$. This cascade will be layered and fed with uranium at higher
enrichment to evaluate the possibility to used them, with few or no tuning, to
produce \gls{HEU}. The characteristics of the default cascade are summarized in
Table \ref{tab:cascade_config}.

@@

\subsection{scenarios}
In the following, cascades can be connected in tandem, where each set of cascade
in parallel is called a ``level``, as illustrated on Figure
\ref{fig:cascade_level}.
Seven different simulations have been simulated, to evaluate the effectiveness
of an enrichment cascade when used outside of its designed scope :
\begin{itemize}
\item one as the reference calculation, with a single centrifuges designed to
    directly produce \gls{HEU} from natural uranium,
\item three calculations (one per misuse model) were default cascade are
    chained to produce \gls{HEU}, without recycling the tails of each cascade,
    the cascade tails are directly sent to the waste,
\item three calculations (one per misuse model) were default cascade are
    chained to produce \gls{HEU}, and the tails of each cascade are recycled, blending
    the tails of one level in the feed of the previous level of
    cascades (see Figure \ref{fig:cascade_level}).
\end{itemize}

\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.45]{flow}
    \caption{Schematic representation of the chained cascades with three levels,
    with the feed, product and the tails flows, respectively in blue, green and
    red. The dashed orange line represent the alternative tails flow when
    tails recycling is considered.}
    \label{fig:cascade_level}
\end{figure}

@@

As it is not possible to assign a fraction of an enrichment cascade, cascade per
level are rounded up for each level but the first one. To the first level is
attributed the remaining available cascades.

@@

As it can be observed on Figures \ref{fig:case1}, \ref{fig:case2} and
\ref{fig:case3}, the different cascade levels required up to $2n+1$ timesteps
before it starts to enrich uranium, where $n$ corresponds to the level number.
This corresponds to the time required for the material to propagate from the
natural uranium source to the n-th level, regardless of the duration of the
timestep (hour, day, month,...). This is also true for the number of timesteps
required to reach the equilibrium assays value in the tails reprocessing cases.
As timestep can be artificially small or large, one will only be considering
equilibrium values flow rates and assays in the following analysis.

>>>>

\subsubsection{Reference}

@@

\subsubsection{Default cascade}

The default cascade is the cascade design for normal civilian enrichment
operation, enriching natural uranium to about $3.5w\%$, with a tails assay close
to $0.3w\%$. This cascade will be layered and fed with uranium at higher
enrichment to evaluate the possibility to use them, with little or no tuning, to
produce \gls{HEU}. The characteristics of the default cascade are summarized in
Table \ref{tab:cascade_config}.

@@

\subsection{Scenarios}
In the following, cascades can be connected in tandem, where each set of cascade
in parallel is called a ``level``, as illustrated on Figure
\ref{fig:cascade_level}.
Seven different simulations have been simulated, to evaluate the effectiveness
of an enrichment cascade when used outside of its designed scope :
\begin{itemize}
\item one as the reference calculation, with a single cascade designed to
    directly produce \gls{HEU} from natural uranium,
\item three calculations (one per misuse model) where default cascade are
    chained to produce \gls{HEU}, without recycling the tails of each cascade
    sending their tails to the waste,
\item three calculations (one per misuse model) where default cascade are
    chained to produce \gls{HEU}, and the tails of each cascade are recycled, blending
    the tails of one level in the feed of the previous level of
    cascades (see Figure \ref{fig:cascade_level}).
\end{itemize}

\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.45]{flow}
    \caption{Schematic representation of the chained cascades with three levels,
    with the feed, product and the tails flows, in blue, green and red,
    respectively. The dashed orange line represent the alternative tails flow when
    tails recycling is considered.}
    \label{fig:cascade_level}
\end{figure}

@@

As it is not possible to assign a fraction of an enrichment cascade, cascade per
level are rounded up for each level but the first one. The remaining available
cascades is attributed to the first level.

@@

As it can be observed on Figures \ref{fig:case1}, \ref{fig:case2} and
\ref{fig:case3}, the different cascade levels required up to $2n+1$ timesteps
before it starts to enrich uranium, where $n$ corresponds to the level number.
This corresponds to the number of timesteps required for the material to
propagate from the natural uranium source to the n-th level, regardless of the
duration of the timestep (hour, day, month,...). This is also true for the
number of timesteps required to reach the equilibrium assays value in the tails
reprocessing cases.  As timestep can be artificially small or large, one will
only be considering equilibrium values flow rates and assays in the following
analysis.

### 2018-06-28 results.tex
finish with gonuke comments

Model       &        &           & A/NR      & A/R       & B/NR      & B/R      & C/NR       & C/R          \\
%Recycling   &        &           & No        & yes       & No        & Yes     & No        & yes        \\
\midrule                                                                                                 
        &            & Feed      & $0.71w\%$ & $1.3w\%$  & $0.71w\%$ & $0.94w\%$ & $0.71w\%$ & $1.66w\%$ \\
Level 0 & Assay      & Product   & $4.13w\%$ & $7.7w\%$  & $4.13w\%$ & $5.43w\%$ & $4.13w\%$ & $9.53w\%$ \\
        &            & Tails      & $0.29w\%$ & $0.5w\%$  & $0.29w\%$ & $0.39w\%$ & $0.29w\%$ & $0.69w\%$  \\
        & Cascades   &           & 26.73     & 26.45     & 26.6      & 25.8      & 26.73     & 26.45      \\
\midrule                                                                                                 
        &            & Feed      & $4.13w\%$ & $11.9w\%$ & $4.13w\%$ & $6.84w\%$ & $4.13w\%$ & $13.0w\%$ \\
Level 1 & Assay      & Product   & $22.8w\%$ & $55.7w\%$ & $20.6w\%$ & $30.7w\%$ & $22.9w\%$ & $69.8w\%$ \\
        &            & Tails      & $1.8w\%$  & $6.6w\%$  & $1.72w\%$ & $2.91w\%$ & $1.81w\%$ & $9.43w\%$  \\
        & Cascades   &           & 2.92       & 3.20      & 2.91      & 3.41     & 2.92      & 3.20       \\
\midrule                                                                                                 
        &            & Feed      & $22.8w\%$ & $55.7w\%$ & $20.6w\%$ & $34.3w\%$ & $22.9w\%$ & $72.6w\%$ \\
Level 2 & Assay      & Product   & $78.5w\%$ & $95.0w\%$ & $61.0w\%$ & $75.8w\%$ & $82.0w\%$ & $98.4w\%$ \\
        &            & Tails      & $4.12w\%$ & $50.9w\%$ & $9.56w\%$ & $17.5w\%$ & $15.7w\%$ & $69.4w\%$ \\
        & Cascades   &           & 0.31      & 0.35      & 0.37      & 0.64      & 0.32      & 0.35      \\
\midrule                                                                                                 
        &            & Feed      & $78.5w\%$ & N.A.      & $61.0w\%$ & $75.8w\%$ & $82.3w\%$ & N.A.      \\
Level 3 & Assay      & Product   & $98.2w\%$ & N.A.      & $90.4w\%$ & $95.0w\%$ & $99.1w\%$ & N.A.      \\
        &            & Tails      & $76.1w\%$ & N.A.      & $79.3w\%$ & $56.1w\%$ & $80.3w\%$ & N.A.      \\
        & Cascades   &           & 0.03      & N.A.      & 0.080     & 0.18      & 0.03      & N.A.      \\
\bottomrule
\end{tabular}
  \label{tab:level}
\end{table}

As illustrated on Figures \ref{sfig:case1_NR}, \ref{sfig:case2_NR} and
\ref{sfig:case3_NR} and summarized on Tab \ref{tab:level}, the different model
don't have the same effect on the cascade behavior. While the models A and C,
allow a quick enrichment gain, with the cascades chaining, respectively
4/23/78/98 and 4/23/82/99, the model B, the enrichment gain is only
4/21/61/90.  Similar effects can be observed when recycling the tails\ldots

@@

As shown on Figures \ref{sfig:case1_R}, \ref{sfig:case2_R} and
\ref{sfig:case3_R}, recycling the tails increases the overall product assay at
all the different levels. As the tails assay of a level $n$ is always higher than
the product assay of the level $n-2$, recycling the tails of level $n$ will
consequently increase the feed assay of level $n-1$ (see Table \ref{tab:level}).
Moreover, with an increased feed assay, tails and product assays increase as
well, increasing de facto the feed assays of respectively cascade levels $n-2$
and $n$...  This effect reduces the number of cascade levels required to reach
\gls{HEU} in case A and C.

@@

As shown Figure \ref{fig:HEU_rate}, recycling increases the final \gls{HEU}
production rate, from $2$ to almost $20$ kg/y when using model A and C, and from
$17$ to $38$ kg/y with the model B. For the reference calculation where all the
available cascades are used within a single large cascade design for direct
\gls{HEU} production, the \gls{HEU} production rate is slightly over $50$ kg/y.

@@

\begin{figure}[h!] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.25]{HEU_prod_rate}
    \caption{Production rate for the different model configuration, with in blue
    the case without tails recycling, in orange with tails recycling, and in green
    the reference one. Where A-B-C represent the model used, and NR-R
    respectively the case without tails recycling and the case with tails recycling}
    \label{fig:HEU_rate}
\end{figure}

>>>>

Model       &        &           & A/NR      & A/R       & B/NR      & B/R      & C/NR       & C/R          \\
%Recycling   &        &           & No        & yes       & No        & Yes     & No        & yes        \\
\midrule                                                                                                 
        &            & Feed      & $0.71w\%$ & $1.3w\%$  & $0.71w\%$ & $0.94w\%$ & $0.71w\%$ & $1.66w\%$ \\
Level 0 & Assay      & Product   & $4.13w\%$ & $7.7w\%$  & $4.13w\%$ & $5.43w\%$ & $4.13w\%$ & $9.53w\%$ \\
        &            & Tails     & $0.29w\%$ & $0.5w\%$  & $0.29w\%$ & $0.39w\%$ & $0.29w\%$ & $0.69w\%$ \\
        & Cascades   &           & 25        & 25        & 25        & 24        & 25        & 25        \\
\midrule                                                                                                 
        &            & Feed      & $4.13w\%$ & $11.9w\%$ & $4.13w\%$ & $6.84w\%$ & $4.13w\%$ & $13.0w\%$ \\
Level 1 & Assay      & Product   & $22.8w\%$ & $55.7w\%$ & $20.6w\%$ & $30.7w\%$ & $22.9w\%$ & $69.8w\%$ \\
        &            & Tails     & $1.8w\%$  & $6.6w\%$  & $1.72w\%$ & $2.91w\%$ & $1.81w\%$ & $9.43w\%$ \\
        & Cascades   &           & 3         & 4         & 3         & 4         & 3         & 4         \\
\midrule                                                                                                 
        &            & Feed      & $22.8w\%$ & $55.7w\%$ & $20.6w\%$ & $34.3w\%$ & $22.9w\%$ & $72.6w\%$ \\
Level 2 & Assay      & Product   & $78.5w\%$ & $95.0w\%$ & $61.0w\%$ & $75.8w\%$ & $82.0w\%$ & $98.4w\%$ \\
        &            & Tails     & $4.12w\%$ & $50.9w\%$ & $9.56w\%$ & $17.5w\%$ & $15.7w\%$ & $69.4w\%$ \\
        & Cascades   &           & 1         & 1         & 1         & 1         & 1         & 1         \\
\midrule                                                                                                 
        &            & Feed      & $78.5w\%$ & N.A.      & $61.0w\%$ & $75.8w\%$ & $82.3w\%$ & N.A.      \\
Level 3 & Assay      & Product   & $98.2w\%$ & N.A.      & $90.4w\%$ & $95.0w\%$ & $99.1w\%$ & N.A.      \\
        &            & Tails     & $76.1w\%$ & N.A.      & $79.3w\%$ & $56.1w\%$ & $80.3w\%$ & N.A.      \\
        & Cascades   &           & 1         & N.A.      & 1         & 1         & 1         & N.A.      \\
\bottomrule
\end{tabular}
  \label{tab:level}
\end{table}

As illustrated on Figures \ref{sfig:case1_NR}, \ref{sfig:case2_NR} and
\ref{sfig:case3_NR} and summarized on Tab \ref{tab:level}, the different model
don't have the same effect on the cascade behavior. While the models A and C,
allow a quick enrichment gain, with the cascades chaining, respectively
4/23/78/98 and 4/23/82/99, the model B, the enrichment gain is only
4/21/61/90.

@@

As shown in Figures \ref{sfig:case1_R}, \ref{sfig:case2_R} and
\ref{sfig:case3_R}, recycling the tails increases the overall product assay at
all the different levels. As the tails assay of a level $n+1$ is always higher than
the product assay of the level $n-1$, recycling the tails of level $n+1$ will
consequently increase the feed assay of level $n$ (see Table \ref{tab:level}).
Moreover, with an increased feed assay, tails and product assays increase as
well, increasing de facto the feed assays of respectively cascade levels $n-1$
and $n+1$, etc.  This effect reduces the number of cascade levels required to reach
\gls{HEU} in case A and C.

@@

As shown in Figure \ref{fig:HEU_rate}, recycling increases the final \gls{HEU}
production rate, from $2$ to almost $20$ kg/y when using model A and C, and from
$17$ to $38$ kg/y with the model B. For the reference calculation where all the
available cascades are used within a single large cascade design for direct
\gls{HEU} production, the \gls{HEU} production rate is slightly over $50$ kg/y.

@@

\begin{figure}[h!] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.25]{HEU_prod_rate}
    \caption{Production rate at equilibrium for the different model
        configurations, the case without tails recycling (blue), with tails
        recycling (orange), and the reference one (green). A-B-C represent
        the model used, and NR-R the case without tails recycling and the case
        with tails recycling, respectively.}
    \label{fig:HEU_rate}
\end{figure}

### 2018-06-28 theory.tex
finish with gonuke comments

\begin{equation}\label{eq:gamma_p}
\gamma(1-N')(P\theta-N) - N'(1-(N'\theta-N)/(-1+\theta))(-1+\theta) = 0
\end{equation}

The only solution allowing product assay to range between 0 and 1 is the
following :
\begin{equation}\label{eq:model_b_sol}
    \frac{\gamma(N + \theta) - N 
            - \sqrt{\gamma^2(N^2 
                            - 2N\theta 
                            + \theta^2) 
                    - \gamma(2N^2 - 2\theta^2 
                            + 2N + 2\theta) 
                    + N^2 + 2N\theta + \theta^2 - 2N - 2\theta + 1} 
            - \theta + 1}
        {2\theta(\gamma - 1)}
\end{equation}
Once the product assay is known, one can trivially determine the tails assay,
$\alpha$ and $\beta$ using equations \eqref{eq:alphabeta} and mass
conservation.

Similarly as model A, because the cut values remain constant, the flows don't
need to be recomputed, and the correct assays, $\alpha$ and $\beta$ are
determined through iterative blending of the product assays of the previous
stage and the tails assay of the next stage using equation
\eqref{eq:model_b_sol}.

>>>>

\begin{equation}\label{eq:gamma_p}
    \theta(\gamma-1)N'^2+((N+\theta)(\gamma-1)+1)N'-N\gamma = 0
\end{equation}

The only solution allowing product assay to range between 0 and 1 is the
following :
\begin{equation}\label{eq:model_b_sol}
    N' = \frac{N+\theta}{2\theta} +
         \frac{1 - \sqrt{\gamma^{2}(N-\theta)^{2}
                         + 2\gamma( N^{2} + N - \theta^{2} + \theta)
                         + (N + \theta + 1)^{2}}}
              {2\theta(\gamma - 1)}
\end{equation}
Once the product assay is known, one can trivially determine the tails assay,
$\alpha$ and $\beta$ using equations \eqref{eq:alphabeta} and mass
conservation.

Similar to model A, because the cut values remain constant, the flows don't
need to be recomputed, and the correct assays, $\alpha$ and $\beta$ are
determined through iterative blending of the product assays of the previous
stage and the tails assay of the next stage using equation
\eqref{eq:model_b_sol}.
